---
title: "part2"
output: html_document
date: "2026-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
data = read_csv("nhanes.csv")
```

```{r}
head(data, 3)
```

# Using part 1's question:

#### Exploration

```{r}
ggplot(data, aes(x=carb_intake)) +
  geom_histogram(fill="darkblue", color="white") +
  labs(title="Distribution of carb intake") +
  theme_minimal()

ggplot(data, aes(x=kcal)) +
  geom_histogram(fill="darkblue", color="white") +
  labs(title="Distribution of kcal") +
  theme_minimal()
```

So kcal becomes our target variable. We can use various predictor variables to serve as "exposures" by assigning thresholds.

```{r}
diet_predictors=select(data, c(1, c(16:21)))
summary(diet_predictors[-1])

density_predictors=select(data, c(1, c(11:14)))
summary(density_predictors[-1])

misc_predictors=select(data, c(1, c(2)))
summary(misc_predictors[-1])

#sdoh_predictors=select(data, c(1, c(16:21)))
#summary(sdoh_predictors[-1])
```

The following might be too advanced for our project...

Let's focus only on carbs for now, as per our part 1 question. [Mayo Clinic](https://www.mayoclinic.org/healthy-lifestyle/nutrition-and-healthy-eating/in-depth/carbohydrates/art-20045705) seems to suggest between 45% to 65% of calories be from carbs for a healthy diet. Also, In general, sugar or starch gives you about 4 calories a gram.

```{r}
target_calories = 2000
target_carb_cals = c(0.45*target_calories, 0.65*target_calories)
mayo_g_cal = 1/4 #1 gram per 4 cals

target_carb_gs = target_carb_cals * mayo_g_cal

#now, going to create a new column for our data table with whether the participant had higher than normal carb intake.

more_carbs = data$carb_intake > target_carb_gs[2]

sum(more_carbs) #counts how many are above our limit
#583, seems enough for data analysis/visualization

data = mutate(data, "more_carbs"=more_carbs)

head(select(data,more_carbs),3)
```

```{r}
custom_labels <- c("TRUE" = "High Carb Intake", "FALSE" = "Low Carb Intake")
ggplot(data, aes(x=kcal)) +
  geom_histogram(fill="darkblue", color="white") +
  facet_wrap(vars(more_carbs), labeller = labeller(more_carbs = custom_labels)) + 
  labs(title="Distribution of kcals faceted by having high carb intake shows unimodal peaks") +
  theme_bw()
```

Now let's look for more general correlations, which are possible points of "exposure".

```{r}
corr_search_data = data

lin_outcome = "kcal"
cor_vars = names(corr_search_data)
cor_vecs = rep(0, length(cor_vars))
for (i in c(1:length(cor_vars))) { #so we run cor() on every variable we have
  if (is.numeric(corr_search_data[[cor_vars[i]]])) {
    cor_vecs[i] = cor(corr_search_data[[lin_outcome]], corr_search_data[[cor_vars[i]]])
  } else {
    cor_vecs[i]=0
  }
}
data.frame(x=cor_vars, corr=cor_vecs) |>
  mutate(r2 = corr ** 2) |>
  filter(x != "id") |>
  filter(x != lin_outcome) |>
  arrange(desc(r2)
```

------------------------------------------------------------------------

#### Filtering Outliers and Testing Normality

We should filter our data to remove outliers (like whatever is beyond 8000), and maybe the tail.

```{r}
data_parameters = summary(data$kcal)
data_parameters
```

```{r}
data |>
  filter(kcal >= 6000) #i cant seem to see anything strange in the other factors
```

```{r}
iqr = data_parameters[[5]] - data_parameters[[2]]
lower_bound = data_parameters[[2]] - (1.5 * iqr)
upper_bound = data_parameters[[5]] + (1.5 * iqr)

filtered_data = data |>
  filter(kcal > lower_bound) |>
  filter(kcal < upper_bound)

print(paste(nrow(data), nrow(filtered_data)))
```

```{r}
ggplot(filtered_data, aes(x=kcal)) +
  geom_histogram(bins=60, fill="darkgreen", color="white") +
  labs(title="Distribution of filtered kcals") +
  theme_bw()

ggplot(filtered_data, aes(sample = kcal)) +
  geom_qq(color="darkgreen", size=0.7) +
  geom_qq_line(color="red") +
  labs(title="QQ Plot of kcal suggests non-uniform distribution") +
  theme_minimal()
```

```{r}
#lets try with numerical data
filtered_kcals = filtered_data$kcal
manual_filt_mean = sum(filtered_kcals) / length(filtered_kcals)
filt_median = quantile(filtered_kcals, 0.5)

print(paste0("Absolute Difference between Mean and Median: ", round(abs(manual_filt_mean - filt_median), 2)))

```

All this suggests that by itself, kcal distributions without filtering for exposures, even with outliers removed, will not have a normal distribution.

------------------------------------------------------------------------
