---
title: "Plots"
format: html
editor: visual
embed-resources: true
---

# Plots regarding BMI/obesity:

The National Health and Nutrition Examination Survey (NHANES) is a questionnaire administered by the US Centers for Disease Control (CDC) that is meant to study diet and health in U.S. adults. Specifically, the NHANES results can be categorized into a dataset where each row represents one adult participant, and columns indicating various health markers such as age, obesity, nutrient consumption, etc. As such, the NHANES dataset provides ways to explore how various diet quality and/or socioeconomic factors affect BMI and obesity.

```{r}
library(readr)
library(dplyr)
library(ggplot2)
data = read_csv("nhanes.csv")
```

```{r}
ggplot(data, aes(x=age, y=bmi)) +
  geom_point(color="darkblue") +
  labs(title="No clear trends between participant age and BMI", x="Participant Age", y="BMI") +
  theme_minimal()
```

```{r}
ggplot(data, aes(x=bmi)) +
  geom_histogram(binwidth=5, fill="darkblue", color="white") +
  labs(title="Distribution of recorded participant BMIs", x="BMI") +
  theme_minimal()

```

```{r}
ggplot(data, aes(x=bmi, fill=sex)) +
  geom_histogram(position="stack", color="white", binwidth=2) +
  labs(title="Distribution of recorded participant BMIs", x="BMI", subtitle="Unimodal, slightly skewed right") +
  theme_minimal()

```

```{r}
ggplot(data, aes(x=health_condition, y=bmi)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title="No clear trend between health condition and BMI", x="Health Condition", y="BMI")
```

```{r}
library(broom)
model = lm(bmi ~ kcal, data=data)
model_int = model[[1]][[1]]
model_slope = model[[1]][[2]]
r2 = glance(model)[[1]][[1]]

ggplot(data, aes(x=kcal, y=bmi)) +
  geom_point() +
  geom_abline(slope = model_slope, intercept = model_int, color="red") +
  theme_minimal() +
  labs(title="There does not seem to be a linear relationship between kcal intake and BMI", subtitle="Line of best fit r^2 = 0.0005", x="Energy Intake (kcal)", y= "BMI")
```

The following graphs will attempt to find a suitable subset to focus visualizations to. They may or may not use code that has not been covered in this class yet.

```{r}
data |>
  ggplot(aes(x=kcal, y=bmi)) +
    geom_point(color="darkblue", size=0.5) +
    facet_wrap(~ health_condition) +
    theme_minimal()
```

```{r}
data |>
  ggplot(aes(x=kcal, y=bmi)) +
    geom_point(color="darkblue", size=0.5) +
    facet_wrap(~ race) +
    theme_minimal()
```

```{r}
data |>
  ggplot(aes(x=health_location)) +
    geom_bar() +
    facet_wrap(~ is_obese) +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r}
data |>
  ggplot(aes(y=kcal, x=bmi)) +
    #geom_bar(position="dodge") +
    geom_point() +
    facet_wrap(~ health_location) +
    theme_minimal()
```

```{r}
data |>
  ggplot(aes(x=bmi)) +
    geom_bar() +
    facet_wrap(~ education) + #college grad?? stronger peak there
    theme_minimal()
```

```{r}
data |> filter(education == "College graduate or above") |>
  filter(health_location == "A doctor's office or health center") |>
  ggplot(aes(x=bmi)) +
    geom_bar() +
    facet_wrap(~ food_security)+
    theme_minimal()
```

```{r}
library(broom)
model = lm(kcal ~ protein_intake, data=data)
model_int = model[[1]][[1]]
model_slope = model[[1]][[2]]
r2 = glance(model)[[1]][[1]]

data |>
  ggplot(aes(x=protein_intake, y=kcal)) +
    geom_point(color="darkblue", size=0.5) +
    geom_abline(slope = model_slope, intercept = model_int, color="red") +
    theme_minimal() +
    labs(title="There is a positive correlation between Protein Intake and Energy Intake",
      subtitle=paste0("y = ", round(model_slope, 2), "x + ", round(model_int, 2), ", r^2 = ", round(r2, 2)),
      x="Day 1 Protein Intake in g",
      y="Day 1 Energy Intake in kcal")
```

Use this cell to test mutations on the dataset and observe corrs to a variable

```{r}
corr_search_data = data |>
  filter(health_condition == "Good" | health_condition == "Very Good" | health_condition == "Excellent") |>
  filter(income_to_poverty >= 1) |>
  filter(income_to_poverty <= 3) |>
  filter(health_location == "A doctor's office or health center") |>
  filter(is_obese == 1) |>
  filter(education == "College graduate or above" | education == "Some college or AA degree") |>
  filter(age <= 33)

corr_search_data = data

lin_outcome = "kcal"
cor_vars = names(corr_search_data)
cor_vecs = rep(0, length(cor_vars))
for (i in c(1:length(cor_vars))) { #so we run cor() on every variable we have
  if (is.numeric(corr_search_data[[cor_vars[i]]])) {
    cor_vecs[i] = cor(corr_search_data[[lin_outcome]], corr_search_data[[cor_vars[i]]])
    #print(paste0("cor(", lin_outcome, ", ", cor_vars[i], "): ", cor_vecs[i]))
  } else {
    cor_vecs[i]=0
  }
}
data.frame(x=cor_vars, corr=cor_vecs) |>
  mutate(r2 = corr ** 2) |>
  filter(x != "id") |>
  #filter(x != "bmi") |>
  #filter(x != "is_obese") |>
  arrange(desc(r2)) #well thats not promising
```

```{r}
students = data |>
  filter(health_condition == "Good" | health_condition == "Very Good" | health_condition == "Excellent") |>
  filter(income_to_poverty >= 1) |>
  filter(income_to_poverty < 4) |>
  filter(health_location == "A doctor's office or health center") |>
  filter(is_obese == 1) |>
  filter(education == "College graduate or above" | education == "Some college or AA degree") |>
  filter(age <= 33)

sm = lm(bmi ~ sodium_den, data=students)
sm_int = sm[[1]][[1]]
sm_slope = sm[[1]][[2]]

print(paste0("# of participants: ", dim(students)[1], " // R^2: ", round(glance(sm)[1][[1]], 2) ))

ggplot(students, aes(x=sodium_den, y=bmi)) +
  geom_point(color="darkblue") +
  labs(title="Obese adults who were under 33 and a student with Good or better health...", subtitle="...who visits a doctor's office or health center with moderate income (1-3x poverty line)", x="Sodium Density in g/1000 kcal", y="BMI") +
  geom_abline(slope = sm_slope, intercept = sm_int, color="red") +
  theme_minimal()
```

------------------------------------------------------------------------

## Extra EDA

```{r}
ggplot(data, aes(x=race, fill=education)) +
  geom_bar(position="dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title="There are many more Non-Hispanic White College Grad participants")
```

```{r}
ggplot(data, aes(x=health_condition)) +
  geom_bar(fill="darkblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title="There are many more participants in Good or Very Good health")
```

```{r}
ggplot(data, aes(x=sex, fill=marriage)) +
  geom_bar() +
  theme_minimal() +
  labs(title="More female participants, and most participants overall are married")
```

```{r}
#somewhat broken
#ggplot(filtered_data, aes(x=income_to_poverty, y=health_location)) +
#  geom_boxplot() +
#  theme_minimal() +
#  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#  labs(title="Income to Poverty Ratio might be correlated to Healthcare Location")
```

```{r}
ggplot(data, aes(x=is_obese, fill=sex)) +
  geom_bar(position="dodge") +
  theme_minimal() +
  labs(title="More people are not obese, and females are more represented in either case", x="Is Obese? (0=no, 1=yes)")
```
